---
title: "Comparison of two Picarro G5310 units"
subtitle: "JKADS5099 vs. JKADS5100"
output: pdf_document
date: '`r paste("First created: 08-Apr-2020. Updated:", format(Sys.Date(),"%d-%b-%Y"))`'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Context

In late 2019, two G5310 units were purchased from Picarro <https://www.picarro.com/products/g5310_gas_concentration_analyzer>. The G5310 provides measurements of carbon monoxide (CO) and nitrous oxide (N~2~O) One was bought by MNF to be deployed on the RV _Investigator_ (JKADS5100); whilst the other (JKADS5099) is to be a test unit based at Aspendale, providing measurements in the context of our Urban Greenhouse Gases work. 
Both units were delivered to Aspendale for side-by-side testing prior to the deployment of JKADS5100 on the RV _Invertigator_. 

The following tests have been performed on the units: 

* 7-point calibration for both CO and N~2~O
* water vapour dependence tests 
* accuracy tests using cylinders 
* side-by-side air sampling 

These tests have uncovered some striking differences between the two units, especially for **water vapour dependence** and **accuracy**. This preliminary report focuses on these aspects of their performance. 

```{r, include = FALSE}
library(tidyverse)
library(lattice)
load(file = "rdata/h2otest_data.RData")
load(file = "rdata/tank_data.RData")
```

## Water vapour dependence tests 

These tests were performed on five cylinders containing various amounts of CO and N~2~O: 

TODO: (insert table of values here - get them from squall, print as a tibble)



```{r}
filter(water, co_flag == "") %>% mutate(co_ratio = co_dry/co_wet) %>% 
  ggplot(aes(x = h2o, y = co_ratio, colour = co_C)) +
  geom_point() +
  labs(x = "Water vapour", 
       y = "dry/wet",
       colour = "CO (ppb)") + 
  facet_grid(~instrument)

filter(water, n2o_flag == "") %>% mutate(n2o_ratio = n2o_dry/n2o_wet) %>% 
  ggplot(aes(x = h2o, y = n2o_ratio, colour = n2o_C)) +
  geom_point() +
  labs(x = "Water vapour (ppm)", 
       y = "dry/wet",
       colour = expression("N"[2]*"O (ppb)")) + 
  facet_grid(~instrument)
```




## Accuracy

...10 cylinders containing various concentrations of CO2 ... 