---
title: "Comparison of two Picarro G5310 units"
subtitle: "JKADS5099 vs. JKADS5100"
output: pdf_document
date: '`r paste("First created: 08-Apr-2020. Updated:", format(Sys.Date(),"%d-%b-%Y"))`'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Context

In late 2019, two G5310 units were purchased from Picarro <https://www.picarro.com/products/g5310_gas_concentration_analyzer>. The G5310 provides measurements of carbon monoxide (CO) and nitrous oxide (N~2~O) One was bought by MNF to be deployed on the RV _Investigator_ (JKADS5100); whilst the other (JKADS5099) is to be a test unit based at Aspendale, providing measurements in the context of our Urban Greenhouse Gases work. 
Both units were delivered to Aspendale for side-by-side testing prior to the deployment of JKADS5100 on the RV _Invertigator_. 

The following tests have been performed on the units: 

* 7-point calibration for both CO and N~2~O
* water vapour dependence tests 
* accuracy tests using cylinders 
* side-by-side air sampling 

These tests have uncovered some striking differences between the two units, especially for **water vapour dependence** and **accuracy**. This preliminary report focuses on these aspects of their performance. 

```{r, include = FALSE}
library(tidyverse)
load(file = "rdata/h2otest_data.RData")
load(file = "rdata/tank_data.RData")
```

## Water vapour dependence tests 

These tests were performed on five? cylinders


```{r}

```
You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Accuracy